---
title: "The Best Cars for New Graduates: Should You Buy a New Car?"
author: "Jeny Kwon"
date: "2/17/2020"
output: pdf_document
---

```{r, message=FALSE, echo=FALSE}
library(googlesheets4)
library(tidyverse)
library(dplyr)
library(knitr)
library(data.table)
library(formattable)
library(stringr)
sheets_deauth()
```

```{r, message=FALSE, echo=FALSE}
honda_fit <- read_sheet("1DBUS0kb9dB0d30toeG6FcGUzsCCMVrPr7lp7dRnyd28")
mini_cooper <- read_sheet("1DBUS0kb9dB0d30toeG6FcGUzsCCMVrPr7lp7dRnyd28", sheet = 2)

honda_crv <- read_sheet("1DBUS0kb9dB0d30toeG6FcGUzsCCMVrPr7lp7dRnyd28", sheet = 3)

kia_forte <- read_sheet("1DBUS0kb9dB0d30toeG6FcGUzsCCMVrPr7lp7dRnyd28", sheet = 4)

mazda3 <- read_sheet("1DBUS0kb9dB0d30toeG6FcGUzsCCMVrPr7lp7dRnyd28", sheet = 5)
```

```{r, message=FALSE, echo=FALSE}
# Summary for Honda Fit
summary_fit <- honda_fit %>% 
  summarize(
    car = "Honda Fit",
    number_of_listings = n(),
    average_price = round(mean(price), digits = 2),
    new_price = 16190,
    savings = round(new_price-mean(price), digits = 2)
  )

# Summary for Mini Cooper
summary_cooper <- mini_cooper %>% 
  summarize(
    car = "Mini Cooper",
    number_of_listings = n(),
    average_price = round(mean(price), digits = 2),
    new_price = 21900,
    savings = round(new_price-mean(price), digits = 2)
  )

# Summary for Honda CR-V
summary_crv <- honda_crv %>% 
  summarize(
    car = "Honda CR-V",
    number_of_listings = n(),
    average_price = round(mean(price), digits = 2),
    new_price = 24350,
    savings = round(new_price-mean(price), digits = 2), 
  )

#Summary for Kia Forte
summary_forte <- kia_forte %>% 
  summarize(
    car = "Kia Forte",
    number_of_listings = n(),
    average_price = round(mean(price), digits = 2),
    new_price = 17790,
    savings = round(new_price-mean(price), digits = 2)
  )

# Summary for Mazda3
summary_mazda <- mazda3 %>% 
  summarize(
    car = "Mazda3",
    number_of_listings = n(),
    average_price = round(mean(price), digits = 2),
    new_price = 21000,
    savings = round(new_price-mean(price), digits = 2)
  )
```

As graduation season approaches, many college seniors are looking to start their new jobs in new cities, but how much does it cost to start life after college? 

There are many expenses that come with relocating to a new city. Expenses include moving costs, storage fees, apartment deposits- which usually consist of security deposit, brokerage fee, first and last months’ rent, and cars. 

Often times, most of these expenses are inevitable, and prices are set by the market. Have furniture you want to move? The fees from moving companies around your area are most likely around the same price. Decided to store furniture until you find an apartment? Most storage facilities have similar rates. Cannot live with your parents? You are going to need to find a new apartment, and although some do not require brokerage fees, you are still bound to market rates for rent.  All these expenses add up quickly and are apparent to soon-to-be graduates. However, graduates can save a significant amount of money by simply buying used cars. 

It turns out that a new graduate can save \$12,464 on average by purchasing a used car instead of a new car after finishing school. This savings amount is actually also greater than the price of an average used car, which is \$7,749. 

```{r, message=FALSE, echo=FALSE}
combined <- rbind(summary_fit, summary_cooper, summary_crv, summary_forte, summary_mazda)

combined_2 <- combined %>% 
  data.frame() %>% 
  mutate(greater_savings = ifelse(savings > average_price, "yes", "no"))

combined_2$average_price <- currency(combined_2$average_price, digits = 0L)
combined_2$new_price <- currency(combined_2$new_price, digits = 0L)
combined_2$savings <- currency(combined_2$savings, digits = 0L)

average_savings <- combined_2 %>% 
  summarize(average_saving = mean(savings),
            average_used = mean(average_price),
            average_new = mean(new_price))
kable(average_savings, col.names = c("Average Saving", "Average Used Car Price", "Average New Car Price"), caption = "Average Savings and Prices")
```

This analysis covers the top 5 cars for recent graduates based on U.S. World & News Report’s (U.S. News) article “15 Best Cars for Recent Graduates in 2019.” ^[ U.S. World & News Report. “15 Best Cars for Recent Graduates in 2019.” https://cars.usnews.com/cars-trucks/best-new-cars-for-recent-college-grads]  The top 5 cars are the Honda Fit, Mini Cooper, Honda CR-V, Kia Forte, and Mazda3. According to U.S. News, they “evaluated the top models for cargo space, safety features, fuel economy, and cabin quality” among other characteristics. They have also focused on affordable cars. 

For each of the five cars, the 2019 new price as well as the average used price were collected. The new price was extracted directly from U.S. News’ article. The used prices were scraped from Craigslist. 


```{r, message=FALSE, echo=FALSE}
used_prices <- ggplot(combined_2, aes(car, average_price)) +
  geom_col(fill="darkred") + 
  geom_text(aes(label = average_price), size=3, vjust = -0.5) +
  labs(title = "Average Price of Used Cars (Craigslist Listings in Western Mass)",
       subtitle = "Top 5 Cars for Recent Graduates",
       caption = "source: Craigslist Listings in Western Massachusetts",
       x ="Car Brand and Make",
       y = "Average Price (in Dollars)") +
  ylim(0, 30000)
used_prices
```

```{r, message=FALSE, echo=FALSE}
new_prices <- ggplot(combined_2, aes(car, new_price)) +
  geom_col(fill="dodgerblue4") + 
  geom_text(aes(label = new_price), size=3, vjust = -0.5) +
  labs(title = "Average Price of New Cars",
       subtitle = "Top 5 Cars for Recent Graduates",
       caption = "source: USNews 15 Best Cars for Recent Graduates in 2019 ",
       x ="Car Brand and Make (2019 Models)",
       y = "Price (in USD)") +
  ylim(0,30000)

          
new_prices
```

For each vehicle, all the Craigslist listings that appeared from the search of each car model were included in the analysis; this includes suggested listings from nearby locations. The nearby location listings were included as vehicles tend to be large purchases, which leads to the geographic market to cover a larger radius for most purchasers. 

In addition, the prices scraped from Craigslist were not always reliable. Some analysis showed that some prices were listed as \$0 and asked interested parties to call for more information. None of the prices were excluded from this analysis; therefore, the \$0 listings could have skewed the average prices. However, the same analysis also showed that some listings had significantly higher prices than the average price. Below is the minimum and maximum analysis data as well as the summary of all the data used. 


```{r, message=FALSE, echo=FALSE}
minmax_fit <- honda_fit %>% 
  summarize(
    car = "Honda Fit",
    min_price = min(price),
    max_price = max(price),
    average_price = mean(price)
  )

minmax_cooper <- mini_cooper %>% 
  summarize(
    car = "Mini Cooper",
    min_price = min(price),
    max_price = max(price),
    average_price = mean(price)
  )

minmax_crv <- honda_crv %>% 
  summarize(
    car = "Honda CR-V",
    min_price = min(price),
    max_price = max(price),
    average_price = mean(price)
  )

minmax_forte<- kia_forte %>% 
  summarize(
    car = "Kia Forte",
    min_price = min(price),
    max_price = max(price),
    average_price = mean(price)
  )

minmax_mazda<- mazda3 %>% 
  summarize(
    car = "Mazda3",
    min_price = min(price),
    max_price = max(price),
    average_price = mean(price)
  )


combined_minmax <- rbind(minmax_fit, minmax_cooper, minmax_crv, minmax_forte, minmax_mazda)

combined_minmax$min_price<- currency(combined_minmax$min_price, digits = 0L)
combined_minmax$max_price<- currency(combined_minmax$max_price, digits = 0L)
combined_minmax$average_price<- currency(combined_minmax$average_price, digits = 0L)

kable(combined_minmax, col.names = c("Car", "Cheapest", "Most Expensve", "Average"), caption ="Min/Max Prices for Each Car Model (Used)")

```


```{r, message=FALSE, echo=FALSE}
# Combined Data for all cars
kable(combined_2, col.names = c("Car",
                           "# of Listings (Craigslist)",
                           "Avg Used Car",
                           "New Car",
                           "Savings",
                           "Saving More Than Used Price?"),
      caption = "Summary of All Data for Each Car Model")
```

Table 3 shows that the average price for a used car is significantly cheaper than for a new car for all the top 5 cars for recent graduates. Savings from purchasing a used car range from $10,000 to almost $16,000. For new graduates, this amount of money can make a big difference for their financial health as they start their new stage in life. The second column shows the number of listings available on Craigslist for each model. This is important as it shows that graduates have many options just in Western Massachusetts. The last column in Table 3 shows whether the amount in savings is actually greater than the average price of a used car, which means that graduates could potentially buy two used cars for the price of one new car. 

```{r, message=FALSE, echo=FALSE}
savings_usd <- ggplot(combined_2, aes(car, savings)) +
  geom_col(fill="darkgreen") + 
  geom_text(aes(label = savings), size=3, vjust = -0.5) +
  labs(title = "Amount in Savings from Buying a Used Car",
                    subtitle = "Top 5 Cars for Recent Graduates",
                    x ="Car Brand and Make",
                    y = "Average Savings (in USD)") +
                    ylim(0,30000)
savings_usd
```
Although buying a used car can save new graduates thousands of dollars, it is also important to note that used cars require more research in the buying process, might not be as reliable, and will probably incur more maintenance costs. New graduates should make sure that they purchase their used cars from a reliable source and get the appropriate car checks before purchase. Furthermore, although this analysis did not focus on the age of used cars, graduates have further flexibility price options depending on the age of the used car.  

When calculating post-graduation moving costs, new graduates should strongly consider purchasing a used car to save $12,498 on average. Having an extra thousands of dollars or not having more debt from a car loan can help new graduates build a stronger financial foundation. 

The data and code behind this analysis can be found here: https://github.com/jnkwon/SDS236_ShortStory 




